# Getting started {#start}

To analyse light-level geolocator data in R we need a couple of R packages as well as functions that allow to run our code. We created a package called _GeoLocTools_ that will be updated frequently. This package contains usefull functions that will be discussed in this manual but can also runs a check on you system (_setupGeolocation()_), detecting packages that are already installed and installs the missing tools directly from CRAN or GitHub.

```{r eval = FALSE}
library(devtools)
# install_github("SLisovski/GeoLocTools") # to be done!
```

While the code above needs to be done only once (and after updates of the `GeoLocTools`package - check for new versions), loading the package into R needs to be done every time and ensures that you can start your analysis streight away:

```{r, message=FALSE, warning=FALSE}
library(GeoLocTools)
setupGeolocation()
``` 

if you see "You are all set!" in your console, the function ran succesfully and you are able to proceed.