# Getting started {#start}

To analyse light-level geolocator data in R we need a couple of R packages as well as functions that allow to run our code. We created a package called _GeoLocTools_ that contains functions that are not nessesarily associated to a certain package put are used in this manual. Importantly the package can also runs a check on you system (function: _setupGeolocation()_), detecting packages that are already on your computer and installs the missing tools directly from CRAN or GitHub.

The package requires _devtools_ (install if nessesary using the _install.packages()_ function). With _devtools_ on your system, you are able to download and built as well as install R packages directly from GitHub (e.g. _GeoLocTools_).

```{r eval = FALSE}
library(devtools)
install_github("SLisovski/GeoLocTools")
```

You are know able to load the package and run the _setupGeolocation()_ function. We recommend to include this line at the beginning of each script you create for a geolocator analysis. Also check (every know and then), if there is a new version of _GeoLicTools_ available. And if that is the case re-install the package using the same code you used for initial installation.

```{r, message=FALSE, warning=FALSE}
library(GeoLocTools)
setupGeolocation()
``` 

if you see "You are all set!" in your console, the function ran succesfully and you are able to proceed.

Amongst other dependencies, the following packages are loaded by this function:

CRAN: geosphere, maptols, raster, MASS, rgdal

GitHub: twGeos, GeoLight, probGLS, SGAT, FLightR

