# Loading data into R {#loadingData}

here are a few functions to read light-level data from different tag types into R. The type of file that you have will determine which function you use. If you have `.lig` from British Antarctic Survey or BioTrack you use the `readLig` found in the `TwGeos` package. For the `GeoLight` and the `SGAT` analyses we will use a tag developed by the Swiss Ornithological Institute that was deployed on a Bee-eather in 2014/2015. For the `FLightR` analysis we will use recordings from an `Integio` logger developped by `Migrate Technology Ltd` that was deployed on a ...


```{r}
ID <- "14SA"
wd <- "Main"
Species <- "MerApi"
AnalCode <- "R1"

lon.calib <- 11.96
lat.calib <- 51.32


raw <- glfTrans(paste0("Data/RawData/", Species, "/", ID, ".glf"))
  names(raw) <- c("Date", "Light")
  raw$Light  <- log(raw$Light+0.0001) + abs(min(log(raw$Light+0.0001)))
head(raw)
```

No matter how the data is read in and what tag type has been used the output needs to have at least two columns (note that names can be slightly different):

1. datetime
2. light

and these columns need to be in a specific format with datetime beeing a `POSIX.` class and light beeing numeric intergers.

```{r}
str(raw)
```

also note, that some ligth values should be log-transformed while some data is already on a log scale (e.g. .lig files).